UMAP reveals cryptic population structure and phenotype heterogeneity in large genomic cohorts

### 2019-08-29 ms/main.tex
minor changes, add some stuff on global structure

\subsection*{Discussion}
By patching together many local genetic neighbourhoods, UMAP visualizations preserve aspects of both local and global structure. Methods such as t-SNE and LargeVis\cite{tang2016visualizing} are similarly motivated, with key differences including in how they measure neighbours and which cost functions are used in optimizing low-dimensional embeddings: pairwise Gaussian approximations and the KL divergence in t-SNE; approximate nearest neighbours and a likelihood function in LargeVis; and fuzzy simplicial set membership and the cross-entropy in UMAP. Further practical and theoretical comparisons can be found in~\cite{2018arXivUMAP}. UMAP comprehensively illustrates genotypic information at fine scales and within the context of global population structure. It is easy to use and fast; given PCA data and a desktop computer, UMAP can be performed in 15 to 25 minutes on a sample of hundreds of thousands of individuals over tens of dimensions. It excels with larger datasets containing individuals with admixed backgrounds, which present discrete and continuous population structure.

@@

\subsubsection*{Caveats}
Changing the PC cutoff, minimum distance, and number of neighbours can change characteristics of the visualizations. Using a minimal number of neighbours (e.g. 5) can result in the formation of disjoint clusters comprised of related family members (fig~\ref{fig:supp_1kgp_families}), and using a low minimum distance can result in clusters becoming more compact, losing visual detail. We used a minimum distance of 0.5 and 15 neighbours; however, default suggested parameters in UMAP (minimum distance of 0.1 and 15 neighbours) generally perform well across datasets.

@@

Both UMAP and t-SNE feature a number of adjustable parameters. Among the parameters that we varied, the number of PCs used in pre-processing of the data has the largest effect for both methods (see figs~\ref{fig:supp_megamontage_pc2_9} and \ref{fig:supp_megamontage_pc10_50}).

>>>>

\subsection*{Discussion}
%By patching together many local genetic neighbourhoods, UMAP visualizations preserve aspects of both local and global structure. Methods such as t-SNE and LargeVis\cite{tang2016visualizing} are similarly motivated, with key differences in how they measure neighbours and which cost functions are used in optimizing low-dimensional embeddings: pairwise Gaussian approximations and the KL divergence in t-SNE; approximate nearest neighbours and a likelihood function in LargeVis; and fuzzy simplicial set membership and the cross-entropy in UMAP. Further practical and theoretical comparisons can be found in~\cite{2018arXivUMAP}. UMAP, however, comprehensively illustrates genotypic information at fine scales and within the context of global population structure. It is easy to use and fast; given PCA data and a desktop computer, UMAP can be performed in 15 to 25 minutes on a sample of hundreds of thousands of individuals over tens of dimensions. It excels with larger datasets containing individuals with admixed backgrounds, which present discrete and continuous population structure.

By patching together many local genetic neighbourhoods, UMAP visualizations preserve aspects of both local and global structure. Methods such as t-SNE and LargeVis\cite{tang2016visualizing} are similarly motivated, with key differences in how they measure neighbours and which cost functions are used in optimizing low-dimensional embeddings: pairwise Gaussian and t-distributed approximations and the KL divergence in t-SNE; approximate nearest neighbours and a likelihood function in LargeVis; and fuzzy simplicial set membership and the cross-entropy in UMAP. Building the global structure in t-SNE requires tuning the \textit{perplexity} parameter, which in practice will need to be different for varying cluster sizes\cite{wattenberg2016how}, while UMAP merges discrete simplicial sets into a coherent global structure. Further practical and theoretical comparisons can be found in~\cite{2018arXivUMAP}.

UMAP comprehensively illustrates genotypic information at fine scales and within the context of global population structure. It is easy to use and fast: given PCA data and a desktop computer, UMAP can be performed in 15 to 25 minutes on a sample of hundreds of thousands of individuals over tens of dimensions. It excels with larger datasets containing individuals with admixed backgrounds, which present discrete and continuous population structure.

@@

\subsubsection*{Caveats}
In contrast to PCA, UMAP has more adjustable parameters. Changing the PC cutoff, minimum distance, and number of neighbours can change characteristics of the visualizations. Using a minimal number of neighbours (e.g. 5) can result in the formation of disjoint clusters comprised of related family members (fig~\ref{fig:supp_1kgp_families}), and using a low minimum distance can result in clusters becoming more compact, losing visual detail. We used a minimum distance of 0.5 and 15 neighbours; however, default suggested parameters in UMAP (minimum distance of 0.1 and 15 neighbours) generally perform well across datasets.

@@

Both UMAP and t-SNE feature a number of adjustable parameters. Among the parameters that we varied, the number of PCs used in pre-processing of the data has the largest effect for both methods (see figs~\ref{fig:supp_megamontage_pc2_9} and \ref{fig:supp_megamontage_pc10_50}). With UMAP, there are other parameters, such as the learning rate and the distance metric; these were left to the default values.
